
```{r, setup, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  tidy = FALSE,
  error = FALSE,
  fig.width = 8,
  fig.height = 8)
```

# simplegraph

> Simple Graph Data Types and Basic Algorithms

[![Linux Build Status](https://travis-ci.org/MangoTheCat/simplegraph.svg?branch=master)](https://travis-ci.org/MangoTheCat/simplegraph)
[![Windows Build status](https://ci.appveyor.com/api/projects/status/github/MangoTheCat/simplegraph?svg=true)](https://ci.appveyor.com/project/gaborcsardi/simplegraph)
[![](http://www.r-pkg.org/badges/version/simplegraph)](http://www.r-pkg.org/pkg/simplegraph)
[![CRAN RStudio mirror downloads](http://cranlogs.r-pkg.org/badges/simplegraph)](http://www.r-pkg.org/pkg/simplegraph)

Simple classic graph algorithms for simple graph classes.
Graphs may possess vertex and edge attributes. 'simplegraph' has
no dependencies and it is writting entirely in R, so it is easy to
install.

## Installation

```{r eval = FALSE}
devtools::install_github("mangothecat/simplegraph")
```

## Usage

```{r}
library(simplegraph)
```

## Creating graphs

`simplegraph` has two ways of creating graphs from data.
The first one is an adjacency list containing vertex names.
This is Euler's famous graph of the bridges of Koenigsberg.
Note that all graphs are directed in `simplegraph`. Undirected graphs
can be emulated with bidirectional edges.

```{r}
bridges <- graph(list(
  "Altstadt-Loebenicht" = c(
    "Kneiphof",
    "Kneiphof",
    "Lomse"
  ),
  "Kneiphof" = c(
    "Altstadt-Loebenicht",
    "Altstadt-Loebenicht",
    "Vorstadt-Haberberg",
    "Vorstadt-Haberberg",
    "Lomse"
  ),
  "Vorstadt-Haberberg" = c(
    "Kneiphof",
    "Kneiphof",
    "Lomse"
  ),
  "Lomse" = c(
    "Altstadt-Loebenicht",
    "Kneiphof",
    "Vorstadt-Haberberg"
  )
))
bridges
```

## Graph metadata

`simplegraph` supports graph metadata on vertices and edges.
To create a graph with metadata, pass two data frames to `graph`,
one for vertices, one for edges.

The first column of the vertex data frame must contain the ids of
the vertices in a character vector.

The first columns of the edge data frame must contain the edges of
the graph, i.e. the tail vertices and the head vertices, given
by the vertex ids.

Here is an example for a graph of actors and movies.

```{r}
vertices <- data.frame(
  stringsAsFactors = FALSE,
  name = c("Tom Hanks", "Cate Blanchett", "Matt Damon", "Kate Winslet",
    "Saving Private Ryan", "Contagion", "The Talented Mr. Ripley"),
  what = c("actor", "actor", "actor", "actor", "movie", "movie", "movie"),
  born = c("1956-07-09", "1966-05-26", "1970-10-08", "1975-10-05",
    NA, NA, NA),
  gender = c("M", "F", "M", "F", NA, NA, NA),
  year = c(NA, NA, NA, NA, 1998, 2011, 1999)
)

edges <- data.frame(
  stringsAsFactors = FALSE,
  actor = c("Tom Hanks", "Cate Blanchett", "Matt Damon", "Matt Damon", 
    "Kate Winslet"),
  movie = c("Saving Private Ryan", "The Talented Mr. Ripley",
    "Saving Private Ryan", "The Talented Mr. Ripley", "Contagion")
)
actors <- graph(vertices, edges)
```  

```{r}
vertex_ids(actors)
vertices(actors)
edges(actors)
```

## Basic queries

Number of vertices and edges:

```{r}
order(bridges)
size(bridges)
```

Adjacenct vertices:

```{r}
adjacent_vertices(bridges)$Lomse
```

This is a graph of function calls from the R package `pkgsnap`
(https://github.com/mangothecat/pkgsnap):

```{r}
funcs <- graph(list(
  drop_internal = character(0),
  get_deps = c("get_description", "parse_deps",
    "%||%", "drop_internal"),
  get_description = "pkg_from_filename",
  parse_deps = "str_trim",
  cran_file = c("get_pkg_type", "r_minor_version", "cran_file"),
  download_urls = c("split_pkg_names_versions", "cran_file"),
  filename_from_url = character(0),
  get_pkg_type = character(0),
  pkg_download = c("dir_exists", "download_urls",
    "filename_from_url", "try_download"),
  r_minor_version = character(0),
  try_download = character(0),
  drop_missing_deps = character(0),
  install_order = character(0),
  restore = c("pkg_download", "drop_missing_deps",
    "install_order", "get_deps"),
  snap = character(0),
  `%||%` = character(0),
  data_frame = character(0),
  dir_exists = character(0),
  pkg_from_filename = character(0),
  split_pkg_names_versions = "data_frame",
  str_trim = character(0)
))
```

List of vertices:

```{r}
vertices(funcs)
```

List of edges:

```{r}
edges(funcs)
```

## Manipulation

Transposing a graph changes the directions of all edges to
the opposite.

```{r}
edges(transpose(funcs))
```

## Walks on graphs

Breadth-first search:

```{r}
bfs(funcs)
```

## Topological sort

```{r}
topological_sort(simplify(funcs))
```

## Multi-graphs and graphs with loop edges

Detecting loop and multiple edges:

```{r}
is_loopy(funcs)
is_multigraph(funcs)
```

Removing loop and multiple edges:

```{r}
is_loopy(remove_loops(funcs))
is_multigraph(remove_multiple(funcs))
```

`simplify` removes both loops and multiple edges, so it creates
a simple graph:

```{r}
is_loopy(simplify(funcs))
is_multigraph(simplify(funcs))
is_simple(simplify(funcs))
```

## License

MIT © [Mango Solutions](https://github.com/mangothecat).
